---
title: "SGPanimation "
output: github_document
---

```{r setup, include=FALSE}

#The following chunk is taken and modified from dgrtwo's github page. It makes is possible for GIFs to render on a github page
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "SGPAnimationFiles/fig-",
  cache.path = "SGPAnimationFiles/cache-"
)
library(knitr)
library(animation)
knitr::opts_chunk$set(echo = TRUE)
opts_knit$set(animation.fun = function(x, options, format = "gif") {
  x = c(knitr:::sans_ext(x), knitr:::file_ext(x))
  fig.num = options$fig.num
  format = sub("^[.]", "", format)
  fig.fname = paste0(sub(paste0(fig.num, "$"), "*", x[1]), 
                     ".", x[2])
  mov.fname = paste0(sub(paste0(fig.num, "$"), "", x[1]), ".", 
                     format)

  # order correctly
  figs <- Sys.glob(fig.fname)
  figs <- figs[order(as.numeric(stringr::str_match(figs, paste0("(\\d+)\\.", x[2]))[, 2]))]

  animation::im.convert(figs, output = mov.fname)

  sprintf("![%s](%s)", options$label, paste0(opts_knit$get("base.url"), mov.fname))
})

opts_chunk$set(cache = TRUE, message = FALSE, warning = FALSE, fig.show = "animate")
```

## Purpose
Purpose of this book is to develop an animation that takes advantage of **gganimate**, **tweenr**, and other R packages. The animation will hopefully provide an intuitive animation for how student growth percentiles are calculated using an approach *similar* to the NYC Department of Education.

**Disclaimer**: I contributed to this book my own personal capacity. The views and content expressed here are my own and do not necessarily represent the views of the NYC Department of Education. 

## Load Packages

```{r packages, include=FALSE}
library(ggplot2)
library(gganimate)
library(ggforce)
library(tweenr)
source('studentAppear.R')
source('plotStudents.R')
source('lmStudents.R')
source('predictStudents.R')
source('findPeers.R')
source('assignSGP.R')
```

## Students of All Backgrounds

*Under Construction - Explanation pending*
```{r appear, echo=FALSE, fig.show='animate', interval=1/15, fig.align='center'}

gganimate(p1, title_frame = FALSE)

```

## Plot Students by Prior Achievement

*Under Construction - Explanation pending*

```{r plot, echo=FALSE, fig.show='animate', interval=1/15}

gganimate(p2, title_frame = FALSE)

```

## Fit a Regression Line

*Under Construction - Explanation pending*

```{r lm, echo=FALSE, fig.show='animate', interval=1/15}

gganimate(p3, title_frame = FALSE)

```

##Estimate Fitted EOY Achievement

*Under Construction - Explanation pending*

```{r predict, echo=FALSE, fig.show='animate', interval=1/15}

gganimate(p4, title_frame = FALSE)

```

##Assign peer groups for each student

*Under Construction - Explanation pending*


```{r findPeers, echo=FALSE, fig.show='animate', interval=1/15}

gganimate(p5, title_frame = FALSE)

```

##Assign Student Growth Percentile

*Under Construction - Explanation pending*



```{r assign, echo=FALSE, fig.show='animate', interval=1/15}

gganimate(p6, title_frame = FALSE)

```

